---
title: "Bad Controls"
bibliography: references.bib
format: html
editor: visual
---

```{r include=FALSE}
library(dagitty)
library(ggdag)
library(ggplot2)

knitr::opts_chunk$set(dev = "png")
```

## A Crash Course in Bad Controls

These are my notes from @cinelli2021crash.

## Good and Bad Controls

Good controls eliminate "omitted variable bias."

Bad controls induce a *non-causal association*.

### You should include all "relevant" variables

Relevant is nuanced here. There are certain variables you do *not* want to include

Bad controls are controls that would be affected by the treatment. They might be thought of as a dependent variable themselves. These conditions are neither necessary or sufficient for deciding whether a control is "good".

## Preliminaries and basic terminology

### Models 1, 2, and 3

```{r echo=FALSE}
dag <- dagitty('dag{
  X [exposure,pos="1,0"]
  Z [pos="2,-1"]
  Y [outcome,pos="3,0"]
  X -> Y
  X <- Z -> Y
}')
ggdag(dag, layout="circle") + theme_void()
```

```{r echo=FALSE}

```

```{r echo=FALSE}
bigger_dag <- dagify(y ~ x + a + b,
                     x ~ a + b,
                     exposure = "x",
                     outcome = "y")
#  automatically searches the paths between the variables labelled exposure and 
#  outcome
dag_paths(bigger_dag)  
ggdag_paths(bigger_dag) + theme_void()
```

```{r}
ggdag_parents(bigger_dag, "x") + theme_void()
```

```{r fig.width=12}
set.seed(123)
x <- rnorm(10)
y <- rnorm(10)

ggplot(data.frame(cbind(x, y)), aes(x = x, y = y)) +
  geom_point(size = 16)
```

```{r}
data.frame(cbind(x, y))
```
